Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2021-08-10 main.tex
One more proof read, couple of tiny changes

Typical cryptographic secrecy involves hiding all transferred data such that external parties in the communication network learn no new information from acquired encryptions. This can be achieved with common symmetric and public-key encryption scheme such as AES \cite{gueronIntelAdvancedEncryption2010} and RSA \cite{rivestMethodObtainingDigital1978}, respectively. In some cases however, partial data leakage or encrypted data processing is required for achieving a desired goal which has led to several homomorphic and functional encryption schemes \cite{paillierPublicKeyCryptosystemsBased1999,shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013,chotardDecentralizedMultiClientFunctional2018} finding uses in such signal processing or localisation tasks. In \cite{alanwarPrOLocResilientLocalization2017}, homomorphic encryption is used to make time-independent model-free location estimates where an estimator does not learn sensor measurements or locations. In \cite{aristovEncryptedMultisensorInformation2018}, similar secrecy is achieved with a linear Kalman filter when a hierarchical sensor network is present. In \cite{shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013}, privacy-preserving aggregation schemes are presented as a means to compute total powergrid usage without disclosing individual contributions, while in \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, centralised and hidden weighted sum aggregations, pWSAc and pWSAh, are introduced as a means for computing local control inputs in a distributed environment without learning individual inputs.

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation and present an implementation that satisfies these requirements, before using it to derive a filter based on the Extended Kalman Filter with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition with different communication assumptions and leakages is given, crucial for its formal security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

As aggregation is typically performed on scalar inputs, we extend our notation to the context of multidimensional estimation data by letting an instance $t_{k,\tau}$ uniquely capture the scalar aggregation during an estimation timestep $k$ for a single element with position index $\tau$. To achieve this in practice, any injective function can be used, such as the concatenation $t_{k,\tau}=k\mathbin\|\tau$. The four algorithms defining our scheme are given as follows.
\begin{description}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose a the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,i\in\{1,\dots,n-1\}$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

\section{Privacy-Preserving Localisation} \label{sec:priv_localisation}
With a concrete scheme meeting the LCAO notion, we can now put forward a localisation filter with communication that can be reformulated to the required protocol. To produce an estimate of the state $\vec{x}_k$, we make use of an algebraic reformulation of the Extended Kalman Filter (EKF), the Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982}, which reduces the filter update step to a single summation. The EIF update step requires the predicted state estimate $\hat{\vec{x}}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ in the information vector and matrix forms
\begin{equation}
    \hat{\vec{y}}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\hat{\vec{x}}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ sensor measurements at time $k$, with measurement models \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\hat{\vec{y}}_{k|k} = \hat{\vec{y}}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\hat{\vec{x}}_{k|k-1}) + \mat{H}_{k,i}\hat{\vec{x}}_{k|k-1})
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\hat{\vec{x}}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $i\in\{1,\dots,n\}$. After then converting the updated information vector and matrix back to state estimate $\hat{\vec{x}}_{k|k}$ and estimate covariance $\mat{P}_{k|k}$, the filter's prediction step can be computed by the navigator locally, using any suitable filter for the known system model \eqref{eqn:system_model}.

@@

which now captures all the private state information in $\hat{\vec{x}}_{k|k-1}$ required at the sensors. The corresponding EIF update steps \eqref{eqn:eif_info_vec_update} and \eqref{eqn:eif_info_mat_update} then become

>>>>

Typical cryptographic secrecy involves hiding all transferred data such that external parties in the communication network learn no new information from acquired encryptions. This can be achieved with common symmetric and public-key encryption scheme such as AES \cite{gueronIntelAdvancedEncryption2010} and RSA \cite{rivestMethodObtainingDigital1978}, respectively. In some cases however, partial data leakage or encrypted data processing is required for achieving a desired goal which has led to several homomorphic and functional encryption schemes \cite{paillierPublicKeyCryptosystemsBased1999,shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013,chotardDecentralizedMultiClientFunctional2018} finding uses in signal processing or localisation tasks. In \cite{alanwarPrOLocResilientLocalization2017}, homomorphic encryption is used to make time-independent model-free location estimates where an estimator does not learn sensor measurements or locations. In \cite{aristovEncryptedMultisensorInformation2018}, similar secrecy is achieved with a linear Kalman filter when a hierarchical sensor network is present. In \cite{shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013}, privacy-preserving aggregation schemes are presented as a means to compute total powergrid usage without disclosing individual contributions, while in \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, centralised and hidden weighted sum aggregations, pWSAc and pWSAh, are introduced as a means for computing local control inputs in a distributed environment without learning individual inputs.

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation and present an implementation that satisfies these requirements, before using it to derive a filter based on the Extended Kalman Filter with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition with different communication assumptions and leakages is given, crucial for its cryptographic security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

As aggregation is typically performed on scalar inputs, we extend our notation to the context of multidimensional estimation data by letting an instance $t_{k,\tau}$ uniquely capture the scalar aggregation during an estimation timestep $k$ for a single element with position index $\tau$. To achieve this in practice, any injective function can be used, such as the concatenation $t_{k,\tau}=k\mathbin\|\tau$. The four algorithms defining our scheme are given as follows.
\begin{description}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,i\in\{1,\dots,n-1\}$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

\section{Privacy-Preserving Localisation} \label{sec:priv_localisation}
With a concrete scheme meeting the LCAO notion, we can now put forward a localisation filter with communication that can be reformulated to the required protocol. To produce an estimate of the state $\vec{x}_k$, we make use of an algebraic reformulation of the Extended Kalman Filter (EKF), the Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982}, which reduces the filter update step to a single summation. The EIF update step requires the predicted state estimate $\hat{\vec{x}}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ in the information vector and matrix forms
\begin{equation}
    \hat{\vec{y}}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\hat{\vec{x}}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ sensor measurements at time $k$, with measurement models \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\hat{\vec{y}}_{k|k} = \hat{\vec{y}}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\hat{\vec{x}}_{k|k-1}) + \mat{H}_{k,i}\hat{\vec{x}}_{k|k-1})
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\hat{\vec{x}}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $i\in\{1,\dots,n\}$. After converting the updated information vector and matrix back to state estimate $\hat{\vec{x}}_{k|k}$ and estimate covariance $\mat{P}_{k|k}$, the filter's prediction step can be computed by the navigator locally using any suitable filter for the known system model \eqref{eqn:system_model}.

@@

which capture all of the private state information in $\hat{\vec{x}}_{k|k-1}$ required at the sensors. The corresponding EIF update steps \eqref{eqn:eif_info_vec_update} and \eqref{eqn:eif_info_mat_update} then become
